---
title: "Untitled"
author: "Carilli"
date: "3/27/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
library(rio)
library(janitor)
library(stringi)
library(hablar)

add_data <- read_csv("additional_data_used.csv")

rank_data <- read_csv("ranked_measure_data_utilized.csv")
```

```{r}
# library(tidyverse)
# 
# file_names <- str_c("excel_files", list.files("excel_files/"), sep = "/")
# 
# tab_names <- lapply(file_names, function(x) excel_sheets(path = x))
# 
# list_names <- lapply
# 
# 
# # list_all <- lapply(file_names, function(x) {
# #   lapply(tab_names, function(y) {read_excel(file = x, sheet = y)}))
# # }
# 
# list_all <- map2(file_names, tab_names, read_excel(file = file_names, sheet = tab_names))
# 
# files <- lapply(file_names, function(y) read_excel(file = y))
# 
# # 
# # tab_names <- lapply(tab_names, function(x) )
# # 
# # list_all <- lapply(tab_names, function(x) read_excel("Full Data.xls", sheet = x))
# # years <- c(2011:2018)
# # for (i in seq_along(years)) {
# #   list_all[[i]]$year = years[i]
# # }
# 
# 
# 
# 
# ```
# 
# Read every spreadsheet every page. 
# 
# ```{r}
# path <- "excel_files/2011 County Health Ranking Virginia Data - v4.xls"
# 
# 
# file_names %>% 
#   map(~excel_sheets(.x)) %>% 
#   set_names() %>% 
#   map(~read_excel, sheet = .x, 
#       path = file_names)
# 
# read_multiple_excel <- function(path){
#   path %>% 
#   excel_sheets() %>% 
#   set_names() %>% 
#   map_df(read_excel, 
#       path = path)
# }
# 
# data <- file_names %>% 
#   map(read_multiple_excel)
#     
# 
# ```
# 
# ```{r}
# library(tidyverse)
# library(readxl)
# library(rio)
# library(janitor)
# 
# file_names <- str_c("excel_files", list.files("excel_files/"), sep = "/")
# 
# tab_names <- lapply(file_names, function(x) excel_sheets(path = x))
# 
# tab_names[1] %>% 
#   map(~str_subset(. ,".Data$")) %>% 
#   set_names() 
# 
# data_list <- map(file_names, import_list) %>% 
#   set_names(str_extract(file_names, "20..")) 
# 
# data_list$`2020`$`Ranked Measure Data` %>% 
#   unname() %>% 
#   row_to_names(1) %>% 
#   clean_names() %>% 
#   glimpse()
# 
# 
#   
# data_list %>% 
#   map_df("Ranked Measure Data") %>% 
#   unname() %>% 
#   row_to_names(1) %>% 
#   clean_names() %>% 
#   glimpse()
#   # names() -> col_names
#   
#  col_names %>% 
#    str_count("na_.") 
#   
#  # map(df, ) 
# # df[[1]]$`Ranked Measure Data`
# ```
# 
# ```{r}
# files <- str_c("excel_files", list.files("excel_files/"), sep = "/")
# 
# map_df(set_names(files), function(file) {
#   file %>% 
#     excel_sheets() %>% 
#     set_names() %>% 
#     map_df(~ read_excel(path = file, sheet = .x), 
#            .id = "sheet")
# },
# .id = "file")
# 
# all_sheets <- 
# files %>% 
#   map(excel_sheets)
# 
# sheets <- c(all_sheets[[1]][[4]], all_sheets[[1]][[5]])
# 
# sheets %>% 
#   set_names()
# 
# files %>% 
#   map(~read_excel(path = .x, sheet = all_sheets[[1]][[4]]))

```


### Final Data Read ranked measure data

```{r}


files <- str_c("excel_files", list.files("excel_files/"), sep = "/")

all_sheets <- 
files %>% 
  map(excel_sheets)

sheets <- c(all_sheets[[1]][[4]], all_sheets[[1]][[5]])

ranked_measure_data_list <- 
  files %>% 
  map(~read_excel(path = .x, sheet = sheets[1])) %>% 
    set_names(str_extract(files, "20..")) %>% 
      map(~ unname(.x)) %>% 
      map(~ row_to_names(.x, row_number = 1)) %>% 
      map(~ clean_names(.x)) %>% 
   map(~ select(.x , 
               -starts_with(c("x95", "x20", "x80", "z_score", "unrel", "sample_size", 
                              "under", "pop", "pcp_r")), 
               -ends_with(c("low", "high", "_2")),
               -state))

years <- c(2011:2020)
for (i in seq_along(years)) {
  ranked_measure_data_list[[i]]$year = years[i]
}

ranked_measure_data <- 
ranked_measure_data_list %>% 
  bind_rows() %>% 
  filter(!is.na(fips)) %>% 
  select(-aggregate_population) %>% 
  hablar::retype() %>% 
  select(1:2, year, everything())

## Use hablar to change types

rm(ranked_measure_data_list, sheets, i, years, all_sheets, files)

ranked_measure_data %>%
  group_by(year) %>% 
  glimpse()


write_csv(ranked_measure_data, "ranked_measure_data.csv")

```

### Read data
```{r}
data <- read_csv("ranked_measure_data.csv")
```

### Find variables with more than 1300 NAs

```{r}
data %>% 
  map(~ sum(length(which(is.na(.x))))) %>% 
  as_tibble() %>% 
  pivot_longer(c(1:190), "var_name") %>% 
  filter(value >= 1300) %>% 
  pull(var_name) ->
  unusable_variables
  
```

### Remove unusuable variables and filter again

```{r}
data %>% 
  select(-unusable_variables) %>% 
  map(~ sum(length(which(is.na(.x))))) %>% 
  as_tibble() %>% 
  pivot_longer(c(1:131), "var_name") %>% 
  filter(value >= 1000) %>% 
  filter(str_detect(var_name,"black")) %>% 
  pull(var_name) %>% 
  c(.,unusable_variables) ->
  unusable_variables
  
```

remove commute variables

```{r}
data %>% 
  select(-unusable_variables) %>% 
  map(~ sum(length(which(is.na(.x))))) %>% 
  as_tibble() %>% 
  pivot_longer(c(1:124), "var_name") %>% 
  arrange(var_name) %>% 
  filter(str_detect(var_name, c("commute", "drive"))) %>% 
  pull(var_name) %>% 
  c(., unusable_variables) -> 
  unusable_variables
```
remove drive variables

```{r}
data %>% 
  select(-unusable_variables) %>% 
  map(~ sum(length(which(is.na(.x))))) %>% 
  as_tibble() %>% 
  pivot_longer(c(1:119), "var_name") %>% 
  filter(str_detect(var_name, "drive")) %>% 
  pull(var_name) %>% 
  c(., unusable_variables) -> 
  unusable_variables
```


```{r}
data %>% 
  select(-unusable_variables) %>% 
  map(~ sum(length(which(is.na(.x))))) %>% 
  as_tibble() %>% 
  pivot_longer(c(1:116), "var_name") %>% 
  filter(str_detect(var_name, c("black", "hispanic"))) %>% 
  pull(var_name) %>% 
  c(., unusable_variables) -> 
  unusable_variables
```

```{r}
unusable_variables <- 
  unusable_variables %>% 
  c(., c("chlamydia_rate", "teen_birth_rate", "pcp_rate", "dentist_rate", "mhp_rate",
         "association_rate","violent_crime_rate", "injury_death_rate", "rates_per_100000",
         "homicide_rate", "mv_mortality_rate"))
```



```{r}
data %>% 
  select(-unusable_variables) %>% 
  map(~ sum(length(which(is.na(.x))))) %>% 
  as_tibble() %>% 
  pivot_longer(c(1:114), "var_name") %>% 
  filter(str_detect(var_name, paste(c("chlamydia_rate", "teen_birth_rate", "pcp_rate", "dentist_rate",
                                      "mhp_rate", "association_rate","violent_crime_rate",
                                      "injury_death_rate", "rates_per_100000",
                                      "homicide_rate", "mv_mortality_rate"),
                                    collapse = "|")
                    )
         ) %>%
pull(var_name) %>% 
  c(., unusable_variables) -> 
  unusable_variables
```

```{r}
data %>% 
  select(-unusable_variables) %>% 
  map(~ sum(length(which(is.na(.x))))) %>% 
  as_tibble() %>% 
  pivot_longer(c(1:104), "var_name") %>% 
  arrange(desc(value))
```


rewrite data

```{r}
data %>% 
  select(-unusable_variables) %>%
  write_csv("ranked_measure_data.csv")
```


More cleaning

```{r}
data <- read_csv("ranked_measure_data.csv")
```

ypll_rate and years_of_,,, are the same variable with different names.

```{r}
data %>% 
  filter(fips != 51000) %>% 
  mutate(ypll_rate = c(ypll_rate[year < 2015], years_of_potential_life_lost_rate[year >= 2015])) %>% 
  select(-years_of_potential_life_lost_rate, years_of_potential_life_lost_rate_white) ->
  data
```

```{r}
data %>% 
  write_csv("ranked_measure_data.csv")
```

```{r}
data <- read_csv("ranked_measure_data.csv")
```

```{r}
data %>% 
  # select(-unusable_variables) %>% 
  map(~ sum(length(which(is.na(.x))))) %>% 
  as_tibble() %>% 
  pivot_longer(everything(), "var_name") %>% 
  arrange(desc(value))
```

```{r}
data %>% 
  map(~ sum(length(which(is.na(.x))))) %>% 
  as_tibble %>% 
  pivot_longer(everything(), "var_name") %>% 
  filter(str_detect(var_name,"limited"))
```

```{r}
data %>% 
  select(percent_lbw)
```

```{r}
data %<>% 
  select(-c(lbw_births, percent_lbw_hispanic, percent_lbw_white)) 
```

```{r}
data %<>% 
  rename(sti_cases = cases)
```


```{r}
data %>% 
  write_csv("ranked_measure_data_utilized.csv")
```


### Variables 

find duplicate variables

```{r}
var_names <- 
ranked_measure_data %>% 
  names()

var_names %>% 
  sort()

var_names[str_detect(var_names, "death")]
```




```{r}
  
  
# ranked_measure_data %>% 
#   select(-starts_with(c("under", "pop", "pcp_r")),
#          -ends_with("_2"),
#          -c("violent_crimes", "psed_num", "aggregate_population")) %>% 
  
# # Find unnecessary variable names.  Applied above.
# variable_names_list <- 
# ranked_measure_data_list %>% 
#   map(~ select(.x , 
#                -starts_with(c("x95", "x20", "x80", "z_score", "unrel", "sample_size")), 
#                -ends_with(c("low", "high")),
#                -state)) %>% 
#   map(~ colnames(.x))
# 
# duplicate_variable_locations <- 
#   c(variable_names_list %>% 
#       map(~ str_sort(.x)) %>% 
#       unlist() %>% 
#       stringi::stri_duplicated()) %>% 
#   which(duplicate_variable_locations == TRUE) 
# 
# 
# var_names <- 
# unlist(names, use.names = FALSE)[
# which(
# unlist(names, use.names = FALSE) %>% 
#   stringi::stri_duplicated() == TRUE)] %>% 
#   unique()
# 
# # ranked_measure_data_list %>% 
# #   map(~ select(.x, var_names))
# 
# variable_names_list %>% 
#   unlist(., use.names = FALSE) %>% 
#   unique() %>% 
#   str_sort()
```


```{r}
library(tidyverse)
library(readxl)
library(rio)
library(janitor)
library(stringi)
library(magrittr)
```


```{r}
data <- read_csv("ranked_measure_data_utilized.csv")
```

```{r}
data %>% 
  map(~ sum(length(which(is.na(.x))))) %>% 
  as_tibble %>% 
  select(ends_with(c("white")), 
         psed_num,
         teen_population,
         number_unemployed,
         labor_force,
         number_children_in_poverty,
         number_dentists,
         number_some_college,
         number_alcohol_impaired_driving_deaths,
         number_driving_deaths,
         cohort_size,
         number_households_with_severe_problems,
         workers, 
         income_ratio
         ) %>% 
  pivot_longer(everything(), "var_name") %>%
  arrange(var_name) %>% 
  pull(var_name) ->
  remove_vars

data %>% 
  select(-remove_vars) %>% 
  write_csv("ranked_measure_data_utilized.csv")
```


```{r}
data %<>% 
  mutate(number_diabetics = c(no_of_diabetics[year == 2011],
           number_diabetics[year != 2011])) %>% 
  select(-no_of_diabetics)
```

```{r}
data %>% 
  select(contains("pm"), year, county)

data %<>% 
  mutate(average_daily_pm25 = c(average_daily_pm25[year <= 2014], 
                                average_daily_pm2_5[year > 2014])) %>% 
  select(-average_daily_pm2_5)

```


```{r}
data %<>% 
  mutate(sti_cases = c(sti_cases[year < 2013],
                             chlamydia_cases[year %in% c(2013, 2014)],
                             number_chlamydia_cases[year >= 2015])) %>%
  select(-contains("chlam"))

write_csv(data,"ranked_measure_data_utilized.csv")  
```


```{r}
data <- 
  read_csv("ranked_measure_data_utilized.csv")
```

```{r}
data %<>% 
  select(-number_single_parent_households, -number_households, -number_household_with_severe_problems)
```

```{r}
data %>% 
  map(~ sum(length(which(is.na(.x))))) %>% 
  as_tibble %>% 
  pivot_longer(everything(), "var_name") %>% 
  arrange(desc(value))
```


### Final Data Additional Measure Data

```{r}
files <- str_c("excel_files", list.files("excel_files/"), sep = "/")

all_sheets <- 
files %>% 
  map(excel_sheets)

sheets <- c(all_sheets[[1]][[4]], all_sheets[[1]][[5]])

years <- c(2011:2020)

files %>% 
  map(~read_excel(path = .x, sheet = sheets[2])) %>% 
    set_names(str_extract(files, "20..")) %>% 
      map(~ unname(.x)) %>% 
      map(~ row_to_names(.x, row_number = 1)) %>% 
      map(~ clean_names(.x)) %>% 
   map(~ select(.x , 
               -starts_with(c("x95", "x20", "x80", "z_score", "unrel", "sample_size")), 
               -ends_with(c("low", "high")),
               -state)
               ) -> add_data

for (i in seq_along(years)) {
  add_data[[i]]$year = years[i]
}

add_data %<>% 
  bind_rows() %>% 
  filter(!is.na(fips),
         !is.na(county)) %>% 
  hablar::retype() %>% 
  select(1:2, year, everything())

write_csv(add_data, "additional_data.csv")
  
```

```{r}
add_data <- read_csv("additional_data.csv")
```

```{r}
add_data %>% 
  map(~ sum(length(which(is.na(.x))))) %>% 
  as_tibble %>% 
  pivot_longer(everything(), "var_name") %>% 
  filter(value == 1336) %>% 
  pull(var_name) ->
  not_useful


add_data %>% 
  select(- not_useful) %>% 
  write_csv("additional_data_used.csv")
```

```{r}
add_data <- read_csv("additional_data_used.csv")

add_data %>% 
  map(~ sum(length(which(is.na(.x))))) %>% 
  as_tibble %>%
  select(contains("pop")) %>% 
  pivot_longer(everything(), "var_name")

add_data %>% 
  select(-contains("population_")) %>% 
  write_csv("additional_data_used.csv")
```

```{r}
add_data <- read_csv("additional_data_used.csv")

add_data %>% 
  map(~ sum(length(which(is.na(.x))))) %>% 
  as_tibble %>%
  # select(contains("pop")) %>% 
  pivot_longer(everything(), "var_name") %>% 
  arrange(var_name)
```

```{r}
add_data %>% 
  map(~ sum(length(which(is.na(.x))))) %>% 
  as_tibble %>%
  select(contains(c("age_adjusted", "child", "drug", "firearm", "hiv",
                    "homicide"))) %>% 
  pivot_longer(everything(), "var_name") %>% 
  arrange(var_name) %>% 
  pull(var_name) -> 
  not_useful

add_data %>% 
  select(-not_useful) %>% 
  write_csv("additional_data_used.csv")
```

```{r}
add_data %>% 
  map(~ sum(length(which(is.na(.x))))) %>% 
  as_tibble %>%
  pivot_longer(everything(), "var_name") %>% 
  arrange(var_name)
```


```{r}
add_data <- read_csv("additional_data_used.csv")

add_data %>% 
  map(~ sum(length(which(is.na(.x))))) %>% 
  as_tibble %>%
  select(contains(c("dentist", "household_", "illiteracy", 
                  "infant_", "liquor", "mph_", 
                  "mortality_", "motor_vehicle"))) %>% 
  pivot_longer(everything(), "var_name") %>% 
  pull(var_name) -> 
  not_useful

add_data %>% 
  select(-not_useful) %>% 
  write_csv("additional_data_used.csv")
```

```{r}
add_data <- read_csv("additional_data_used.csv")

add_data %>% 
  map(~ sum(length(which(is.na(.x))))) %>% 
  as_tibble %>%
  select(contains("number"), -contains(c("food", "zip", "insured"))) %>% 
  pivot_longer(everything(), "var_name") %>% 
  pull(var_name) ->
  not_useful

add_data %>% 
  select(-not_useful) %>% 
  write_csv("additional_data_used.csv")
```

```{r}
add_data <- read_csv("additional_data_used.csv")
rank_data <- read_csv("ranked_measure_data_utilized.csv")

add_data %>% 
  map(~ sum(length(which(is.na(.x))))) %>% 
  as_tibble() %>% 
  pivot_longer(everything(), "var_name")
```

```{r}
rank_data %>% 
  names()

add_data %>% 
  names()
```

```{r}
add_data <- read_csv("additional_data_used.csv")
rank_data <- read_csv("ranked_measure_data_utilized.csv")
```

```{r}
add_data %>% 
  map(~ sum(length(which(is.na(.x))))) %>% 
  as_tibble() %>% 
  select(contains("death")) %>% 
  pivot_longer(everything(), "var_name") %>% 
  pull(var_name) ->
  not_useful

add_data %>% 
  select(-not_useful) %>% 
  write_csv("additional_data_used.csv")
```

```{r}
number_na <- function(df){
  df %>% 
    map(~ sum(length(which(is.na(.x))))) %>% 
    as_tibble %>% 
    pivot_longer(everything(), "var_name") %>% 
    arrange(desc(value))
}

number_na(add_data)
number_na(rank_data)
```

```{r}
add_data %>% 
  select(contains("number")) %>% 
  names() -> 
  not_useful

not_useful <- 
  not_useful[-c(6, 9, 21)]

add_data %>% 
  select(-not_useful) %>% 
  write_csv("additional_data_used.csv")

add_data <- read_csv("additional_data_used.csv")
```

```{r}
add_data %>% 
  names()
```

```{r}
not_useful <- c("households",
"no_drive_alone",
"workers",
"costs",
"percent_couldnt_access",
"non_hispanic_white",
"mhp_rate",
"mhp_ratio",
"percent_uninsured_2",
"other_pcp_rate",
"other_pcp_ratio",
"other_pcp_rate_2",
"other_pcp_ratio_2",
"percent_18",
"percent_65_and_over",
"percent_african_american",
"percent_american_indian_alaskan_native",
"percent_native_hawaiian_other_pacific_islander",
"percent_hispanic",
"percent_non_hispanic_white",
"percent_female",
"percent_rural",
"percent_disconnected_youth",
"non_petitioned_cases")

add_data %>% 
  select(-not_useful) %>% 
  write_csv("additional_data_used.csv")

add_data <- read_csv("additional_data_used.csv")
```
```{r}
add_data %>% 
  names() %>% 
  str_extract("percent_.*") %>% 
  sort()
rank_data %>% 
  names() %>% 
  str_extract("percent_.*") %>% 
  sort()
```

```{r}
tibble(add_data$percent_uninsured, rank_data$percent_uninsured)
```

```{r}
not_useful <- c("number_low_birthweight_births",
"number_single_parent_households",
"number_uninsured",
"number_with_access",
"number_workers")

rank_data %>% 
  select(-not_useful) %>% 
  write_csv("ranked_measure_data_utilized.csv")
```

```{r}
not_useful <- c("percent_binge_drinking", "percent_uninsured")

add_data %>% 
  select(-not_useful) %>% 
  write_csv("additional_data_used.csv")
```

```{r}
add_data %>% names()

rank_data %>% names()
```

```{r}
rank_data%>% 
  map(~ sum(length(which(is.na(.x))))) %>% 
  as_tibble %>% 
  select(contains("food")) %>% 
  pivot_longer(everything(), "var_name") 
    

add_data %>% 
  map(~ sum(length(which(is.na(.x))))) %>% 
  as_tibble %>% 
  select(contains("food")) %>% 
  pivot_longer(everything(), "var_name") 
```

```{r}
rank_data %>% 
  select(county, year, contains("food")) %>% 
  left_join(add_data %>% 
              select(county, year, contains("food"))) 
  
```

```{r}
rank_data %>% 
  select(county, year, contains("zip")) %>% 
  left_join(add_data %>% 
              select(county, year, contains("zip"))) 
```

```{r}
rank_data %>% 
  select(county, year, number_zip_codes) 

add_data %>% 
  select(county, year, number_zip_codes)
  
```

```{r}
add_data %>% 
  left_join(rank_data, by = c("fips", "county", "year")) %>% 
  write_csv("desert_data.csv")

```

```{r}
data <- read_csv("desert_data.csv")
```

```{r}
data %>% 
  lm(percent_obese ~ percent_limited_access_to_healthy_foods, .) %>% 
  summary()
```


```{r}

```

